\chapter{Conclusions}
\label{ch:conclusions}
In this thesis, we have carried out a formal comparison of state-of-the-art
algorithms. From these, we have chosen three algorithms with the optimality
criterion in mind, and we also wanted to keep the implementation light so that
they could be easily maintained. We have shown their implementation providing
an accurate description of how they work. \newline
Finally, two sets of tests have been devised, one based on random scenarios
regarding the \acrs{MAPF} problem, and one based on scenarios taken from a real
warehouse. The results of the tests show that the algorithms, as they are, need
to be improved. Indeed, they fail to provide a solution in many cases, with
\acrs{CP} taking too much time and \acrs{CBS} failing to find a possible
solution when using spanning trees. As stated, sub-optimal and more efficient
versions of the proposed algorithms exist, especially for \acrs{CBS}, but they
imply major revisions of the code in many cases introducing heuristics and
sub-optimality, which was outside the scope of the project \newline
The future work will focus on improving this existing algorithms in order to
handle corner cases that are difficult to solve. Moreover, we would like to
explore the combination of \acrs{CBS} and \acrs{CP}. Indeed, \acrs{CP} fails to
find solutions especially for time and memory limitations, but these problems
could be greatly reduced when the constraints are thought for \acrs{sapf} and
not for a \acrs{mapf} problem. Solving only a small area of the graph with
\acrs{CP} for a certain agent given some constraint could speed up the
\acrs{CBS} process a lot and provide an optimal solution. 
